Self Regulation Ontology Retest Analyses

### 2018-10-15 code/SRO_Retest_Analyses.Rmd
refactoring

demog_rel_df %>%
  mutate(timed = ifelse(var %in% timed_demogs, 1, 0)) %>%
  ggplot(aes(icc, fill = factor(timed)))+
  geom_histogram(position = "identity", alpha = 0.7, color = NA)+
  xlab("ICC")+
  ylab("Count")+
  scale_fill_manual(values = c("grey", "grey25"))+
  theme(legend.position = "none")

ggsave('DemogRelDist.jpg', device = "jpeg", path = "/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/figures/", width = 6, height = 5, units = "in", dpi=500)
```

```{r echo=FALSE}
rm(test_demog, retest_demog, demog_rel_df, numeric_cols)
```

@@

Due to our data collection strategy we did not strictly control for the delay between the two measurements as standard psychometrics studies measuring retest reliability might.

An individual difference measure would preferably remain stable regardless of the delay between multiple measurements.

Since we only had two measurements we could not test directly whether a measure becomes less reliable depending on the delay between the two time points (The average number of days between two measurements would be the same for all measures since reliability is a measure level metric while days between completion a subject level one).

So if you regress the average difference for each measure on average delay between two measurements you are regressing a vector with varying numbers on a vector of same values, like a t test asking if the mean of the varying column, in this case the average difference score, is different than the unique value in the single value column, i.e. the average delay between two time points (Note that this should be true in theory but in practice since for each measure there might subjects for whom the dv could not be calculated there might be >1 unique values for the average delay between the two time points). This analysis is not meaningful.

@@

Make data frame with difference between two scores for each measure for each subject. Since the scores for different measures are on different scales for comparability the difference scores are normalized (ie demeaned and dividied by the sd of the difference score distribution.) Note that the variance of the difference score distribution accounts for the variance in both time points by summing them. Normalization equates the means of each difference score distribution to 0 which would mask any meaningful change between the two time points but the analysis here does not interpret the mean of the difference score distributions but is interested in its relation to the days between completion. We check if the variables show systematic differences between the two points later.

Here we check if the difference is larger the longer the delay

```{r warning=FALSE}
numeric_cols = c()

for(i in 1:length(names(test_data))){
  if(is.numeric(test_data[,i]) & names(test_data)[i] %in% names(retest_data)){
    numeric_cols <- c(numeric_cols, names(test_data)[i])
  }
}

@@

```{r echo=FALSE}
rm(tmp, i)
```

@@

```{r}
t1_2_difference %>%
  ggplot()+
  geom_smooth(aes(as.numeric(days_btw), abs(difference), group=factor(dv)), method='lm', se=FALSE, color = 'grey', alpha = 0.5)+
  geom_smooth(aes(as.numeric(days_btw), abs(difference)), method='lm', color = "black", se=FALSE)+
  theme(legend.title = element_blank())+
  xlab('Days between completion')+
  ylab('Scaled difference score')

ggsave('DaysBtwEffect.jpg', device = "jpeg", path = "/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/figures/", width = 6, height = 4, units = "in", limitsize = FALSE, dpi = 500)
```

@@

summary(tmp <- MCMCglmm(abs(difference) ~ scale(days_btw), random = ~us(scale(days_btw)):dv, data=t1_2_difference))
```

>>>>

demog_rel_df %>%
  mutate(timed = ifelse(dv %in% timed_demogs, 1, 0)) %>%
  ggplot(aes(icc, fill = factor(timed)))+
  geom_histogram(position = "identity", alpha = 0.7, color = NA)+
  xlab("ICC")+
  ylab("Count")+
  scale_fill_manual(values = c("grey", "grey25"))+
  theme(legend.position = "none")
```

```{r echo=FALSE}
rm(test_demog, retest_demog, demog_rel_df, timed_demogs)
```

@@

@@

To check this we make data frame with difference between two scores for each measure for each subject. Since the scores for different measures are on different scales for comparability the difference scores are normalized (ie demeaned and dividied by the sd of the difference score distribution.) Note that the variance of the difference score distribution accounts for the variance in both time points by summing them. Normalization equates the means of each difference score distribution to 0 which would mask any meaningful change between the two time points but the analysis here does not interpret the mean of the difference score distributions but is interested in its relation to the days between completion. We check if the variables show systematic differences between the two points later.

Here we check if the difference is larger the longer the delay.

```{r warning=FALSE}
numeric_cols = get_numeric_cols(test_data)

@@

```{r echo=FALSE}
rm(tmp, i, numeric_cols)
```

@@

```{r}
t1_2_difference %>%
  ggplot()+
  geom_smooth(aes(as.numeric(days_btw), abs(difference), group=factor(dv)), method='lm', se=FALSE, color = 'grey', alpha = 0.5)+
  geom_smooth(aes(as.numeric(days_btw), abs(difference)), method='lm', color = "black", se=FALSE)+
  theme(legend.title = element_blank())+
  xlab('Days between completion')+
  ylab('Scaled difference score')
```

@@

summary(tmp <- MCMCglmm(abs(difference) ~ scale(days_btw), random = ~us(scale(days_btw)):dv, data=t1_2_difference, verbose=FALSE))
```
